New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification

### 2025-06-20 intro.tex
Added references to the intro

Distribution shift is an important problem facing machine-learning practitioners  [\cite{zhang2023dive, sugiyama2007covariate}].  One particular way a distribution shift may manifest in classification problems  is when the class priors change between the training set used to train a model, and the data distribution in deployment -- the so called label shift [\cite{vsipka2022hitchhiker,lipton2018detecting}]. To make a concrete example, consider a driver assistance system recognizing traffic signs. Suppose a speed limit sign was maliciously altered to indicate 80mph limit instead of 30mph. To avoid being fooled by the alteration, the system could utilize priors. The distribution of traffic signs will change markedly when the car leaves highway and enters streets in a city. It is very unlikely to encounter 80mph limit in a city setting. By taking into account the city priors, the system is likely to make the proper classification. In general a well-designed classification system will be able to improve its performance when supplied with a new set of class priors that correspond to the true distribution of classes.

@@

There is a subtle issue with this approach  when handling probabilistic classifiers arising in one-vs-one classification framework.   One-vs-one classification is a two-step framework to approach  multi-class classification. In the first step it reduces the multiclass problem to a series of binary classification problems. In the second step it deduces the multi-class decision by aggregating decisions for the binary problems. Since the procedure is two-step, there is a possibility to apply adjustment of priors either to the probabilistic distributions obtained after the first step, or after the second step. It is not clear what should be the preferred course of action and this problem motivates our paper.

>>>>

Distribution shift is an important problem facing machine-learning practitioners  [\cite{zhang2023dive, sugiyama2007covariate}].  Using probabilistic formulation, we say the distribution shift occurs if there is a change  in the joint probability distribution $p(X,Y)$ of covariates $X$ with the dependent variable $Y$  between training and testing. Noting the factorizations 
$$
p(X,Y) = p(Y\mid X)p(X) = p(X \mid Y )p(Y)
$$
one may distinguish the following subcategories of distribution shift:
\begin{itemize}
	\item covariate shift, if $p(X)$ changes, but $p(Y\mid X)$ is unchanged,
	\item label shift, if $p(Y)$ changes, but $P(X \mid Y)$ remains the same,
	\item concept shift, if $p(Y \mid X)$ itself changes.
\end{itemize}

We will be concerned with label shift in classification problems, where $p(Y)$ is just the prior on classes.   To provide a concrete example, consider a driver assistance system recognizing traffic signs. Suppose a speed limit sign was maliciously altered to indicate 80mph limit instead of 30mph. To avoid being fooled by the alteration, the system could utilize priors. The distribution of traffic signs will change markedly when the car leaves highway and enters streets in a city. It is very unlikely to encounter 80mph limit in a city setting. By taking into account the city priors, the system is likely to make the proper classification. In general a well-designed classification system will be able to improve its performance when supplied with a new set of class priors that correspond to the true distribution of classes.

@@

There is a subtle issue with this approach  when handling probabilistic classifiers arising in one-vs-one classification framework.   One-vs-one classification is a two-step framework to solve  multi-class classification problems. In the first step it reduces the multiclass problem to a series of binary classification problems. In the second step it deduces the multi-class decision by aggregating decisions for the binary problems. Since the procedure is two-step, there is a possibility to apply the adjustment of priors either to the probabilistic distributions obtained after the first step, or after the second step. It is not clear what should be the preferred course of action and this problem motivates our paper.

Some aspects of label shift were studied in earlier works including test time adaptation
[\cite{vsipka2022hitchhiker}], shift detection and quantification [\cite{lipton2018detecting}], statistical properties of label shift estimators [\cite{garg2020unified}], and Fisher consistency of estimators of priors [\cite{tasche2017fisher}].
